---
title: "Rmd Exercise Template"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data

Load the gapminder dataset

```{r chunk name}
# Here is where your R code goes. ``` opens the code chunk, followed by {r}. Then ``` closes the chunk. When you 'knit' the file, the codes will evaluate.
gapminder <- read.csv(here::here("data/gapminder5.csv"), stringsAsFactors=FALSE)
```

## If Statements

Check if there are any records after 2002 in the data set. Return nothing found if there's no such record.
```{r}
check <- 2002
if(any(gapminder$year==check)){
    print("There exists year after 2002.")
}else{
    print("Nothing found.")
}
```
##Loop and If Statements
```{r}
mean_for_all <- mean(gapminder$pop)

for (i in unique(gapminder$country)){
    country_mean <- mean(gapminder$pop[gapminder$country==i])
    if( country_mean < mean_for_all){
        print(paste(i,"has lower population mean."))
    }
}
```
```{r}
low <- 50
high <- 70
for (i in unique(gapminder$continent)){
    continent_mean <- mean(gapminder$lifeExp[gapminder$continent==i])
    if(continent_mean < low){
        print(paste(i,"has lifeExp lower than 50."))
    }
    else if (low < continent_mean & continent_mean < high){
        print(paste(i,"has lifeExp between 50 and 70."))
    }
    else
        print(paste(i,"has lifeExp higher than 70."))
}
```

## Including Plots

You can also embed plots, for example:

```{r another chunk name}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
