---
title: "Day 1 Part 2 Exercisr----Zekai, ZHU"
date: "`r Sys.Date()`"
output: html_document
params:
    answers: TRUE
---

```{r, echo=FALSE, eval=TRUE}
answers<-params$answers
```

```{r global_options, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo=answers, eval=answers,
                      warning = FALSE, message = FALSE,
                      cache = FALSE, tidy = FALSE)
```

## Load the data
Load the 'gapminder' dataset.

```{asis}
### Answer
```

```{r 1}
gapminder <- read.csv("data/gapminder5.csv")
```

## If Statement
Use an if() statement to print a suitable message reporting whether there are any records from 2002 in the gapminder dataset. Now do the same for 2012.

Hint: use the any function.
 
```{asis}
### Answer
```

```{r 2}
year <- 2012
if(any(gapminder$year == year)){
    print(paste("Records found for the year:", year))
}else{
    print(paste("No records found for the year:", year))
}
    
```

## Loop and If Statements
Write a script that finds the mean life expectancy by country for countries whose population is below the mean for the dataset

```{asis}
### Answer
```

```{r 3}
threshold <- mean(gapminder$lifeExp)   #Overall mean life expectancy 

countries <- unique(gapminder$country)   

for (i in countries) {
    mean_le <- mean(gapminder$lifeExp[gapminder$country == i], 
                    na.rm = T)  
    if (mean_le < threshold){
        print(paste0(i, "'s mean life expectancy is: ", mean_le ))
    }
}
```

## Loop and If Statements
Write a script that loops through the gapminder data by continent and prints out whether the mean life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{asis}
### Answer
```

```{r 4}
conts <- unique(gapminder$continent)
lower_bound <- 50
higher_bound <- 70

for (i in conts) {
    mean_le <- mean(gapminder$lifeExp[gapminder$continent == i], 
                    na.rm = T)  
    if (mean_le < lower_bound){
        print(paste0(i, "'s mean life expectancy is less than 50 which is ", mean_le ))
    }else if (mean_le > lower_bound & mean_le < higher_bound){
        print(paste0(i, "'s mean life expectancy is between 50 and 70 which is ", mean_le ))
    }else{
        print(paste0(i, "'s mean life expectancy is greater an 70 which is ", mean_le ))
    }
}
```

##Exercise: Write Functions
Create a function that given a data frame will print the name of each column and the class of data it contains. Use the gapminder dataset. 

Hint: Use mode() or class() to get the class of the data in each column. Remember that names() or colnames() returns the name of the columns in a dataset.

```{asis}
### Answer
```

```{r 5}
report_name_class <- 
    function(df) {
        for(i in names(df)){
            print(paste("Column names:", i, "; Class:", class(df[,i])))
        }
    }

report_name_class(gapminder)

```

##Exercise: Write Functions
Create a function that given a vector will print the mean and the standard deviation of a vector, it will optionally also print the median. 

Hint: include an argument that takes a boolean (TRUE/FALSE) operator and then include an if statement.

```{asis}
### Answer
```

```{r 6}
vector_mean_sd_med <- 
    function(v,include_median) {
        print(paste("Mean:", mean(v), "; STD:", sd(v)))
        if (include_median == TRUE){                                 
            print(paste("Median:", median(v)))
        }
    }
    
x <- gapminder$lifeExp
vector_mean_sd_med(x, TRUE)

```

## Analyzing the relationship

What is the relationship between GDP per capita and life expectancy? Does this relationship change over time? (Hint: Use the natural log of both variables.)

```{asis}
### Answer
```

```{r 7}
years <- unique(gapminder$year)

for(i in years){
     plot(log(gapminder$lifeExp[gapminder$year==i]) ~ log(gapminder$gdpPercap[gapminder$year==i]),
                 main=paste0("Relationship between life expectancy and GDP per capita in ",i),
                 ylab="Life expectancy", xlab="GDP per capita")
    abline(h = mean(gapminder$lifeExp[gapminder$year == i]))
}

```
##Does the relationship between GDP per capita and life expectacy vary by continent? Make sure you divide the Americas into North and South America.

```{asis}
### Answer
```

```{r 7}
```

