---
title: "Day 1 Part 2 Exercises"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Data
```{r}
dat = read.csv(here::here("data/gapminder5.csv"), stringsAsFactors=FALSE)
```

# If Statement
```{r}
year_present = function(year) {
    if (year %in% unique(dat$year)) {
        print(paste("The year", year, "is represented in the data")) }
    else {
        print(paste("The year", year, "is not represented in the data"))
    }
}

year_present("2002")
year_present("2012")
```

# Loop and If Statements
```{r}
mean_life_exp = mean(dat$lifeExp, na.rm=T)
for (country in unique(dat$country)) {
    country_mean = mean(dat$lifeExp[dat$country == country], na.rm=T)
    if (country_mean < mean_life_exp) {
        print(paste(country, ":", country_mean))}
}
```

```{r}
for (continent in unique(dat$continent)) {
    mean_life_exp = mean(dat$lifeExp[dat$continent==continent], na.rm=T)
    if (mean_life_exp<50){
        print(paste("The life expectancy of", continent, "is under 50"))}
    else if (mean_life_exp>70) {
        print(paste("The life expectancy of", continent, "is greater than 70"))}
    else {
        print(paste("The life expectancy of", continent, "is between 50 and 70"))}
}
```

## Write Functions
```{r}
describe_df = function(df) {
    for (i in names(df)) {
        print(paste(i, ":", class(i)))
    }
}
describe_df(dat)
```

```{r}
describe_vector = function(vector, median=FALSE) {
    print(paste("The mean is", mean(vector)))
    print(paste("The standard deviation is", sd(vector)))
    
    if (median==T) {
        print(paste("The median is", median(vector)))
    }
}

describe_vector(c(1,2,3,4,5,6,78,8,7,6,5,4,3,2,2,4,54), median=T)
```

```{r}
plot(dat$gdpPercap, dat$lifeExp)
cor(dat$gdpPercap, dat$lifeExp)
```

There is clearly a positive correlation between life expectancy and GDP per capita, though the relationship is not linear!

```{r}
for (continent in unique(dat$continent)) {
    print(paste("The correlation between GDP per capita and life expectancy in ", continent, "is", cor(dat$gdpPercap[dat$continent==continent], dat$lifeExp[dat$continent==continent])))
}
```

The relationship between GDP per capita and life expectancy definitely changes by continent. The Americas are not split and there is not enough time to go through country by country!