---
title: "ggplot_rmd"
author: "Xiaoling Zhang"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ggplot
### Exercise 1: Scatterplot

Using the gapminder07 data, create a scatterplot of the natural log of `gdpPercap` as a function of the natural log of `pop`. Give it a title and axis labels.

Remember, you will need three functions: `ggplot()`, `geom_point()`, and `labs()`.

```{r}
library(ggplot2)
ggplot(data = gapminder07) + geom_point(mapping = aes(x = log(pop), y = log(gdpPercap))) + labs(title = 'Relationship between the natural log of gdpPercap and the natural log of pop', x = 'log gdpPercap', y = 'log pop')
```

### Exercise 2: Hydro power generated over time

Task: Plot a column chart hydroelectric power generated over time.

Hint: There are two types of hydroelectric sources in the data: `large_hydro` and `small_hydro`.

```{r}
long_gen %>% filter(source == 'large_hydro' | source == 'small_hydro') %>% group_by(datetime) %>% summarise(output = sum(output)) %>% ggplot() + geom_col(aes(x = datetime, y = output))
```


```{r}
long_gen %>%
    group_by(datetime, source) %>% filter(grepl('hydro', source))
```


### Exercise 3: Total output per source

Task: Create a column chart that shows the total output per source.

- Change the color of the columns to `"darkred"`.
- Add a horizontal line indicating the mean output across all sources. Use the cheatsheet to identify the `geom` function that you need.
- Add a meaningful title and axis labels using `labs()`.
```{r}
long_merged_energy %>%
    group_by(source) %>%
    summarise(output = sum(output)) %>%
    ggplot() + geom_col(aes(x = source, y = output), fill = 'darkred') + geom_hline(aes(yintercept = mean(output))) + geom_hline(aes(yintercept = median(output)), linetype = "dashed") + labs(title = "Total energy output", y = 'output', x = 'source')
```


### Exercise 4: Colors and fill

Task: Create a line plot that compares generation of wind, solar, and geothermal energy over time.

Bonus: Set the line size to 1.5.
```{r}
long_merged_energy %>% filter(source == 'wind' | source == 'solar' | source == 'geothermal') %>%
    ggplot() + 
    geom_line(aes(x = datetime, y = output, group = source, col = source), size=1.5)
```

